Adaptive broadcast tree for multicores

### 2013-06-25 paper.tex
More measurements

\documentclass{article}
\usepackage{url,color,xspace,verbatim,subfig,ctable,multirow,listings}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{rotating}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{graphics}
\usepackage{enumitem}
\usepackage{times}
\usepackage{amssymb}
\usepackage[colorlinks=true]{hyperref}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[toc,page]{appendix}
% ==================================================

@@

% tikz stuff
\usepackage{tikz}
\usepackage{pgfplots}
% configuration
\usetikzlibrary{shapes,positioning,calc,snakes,arrows,shapes,fit,backgrounds}
\pgfplotsset{width=.9\linewidth}

@@

@@

\begin{description}
\item[cost breakdown] (quantitative) of message send cost:
  dominating propagation time vs.\ dominating send and receive time on
  multicores (per-node). Our model is currently limited to only the
  breakdown of the cost for sending messages. Maybe we do not strictly
  need to know about propagation time, as it is negligible independent
  of were messages are sent. Relevant seems to be the number of hops
  rather than the latency expressed in a concrete time.
\item[bounded delivery]
\item[reliability] (loss, link failure, in-order delivery, phantoms,
  corruption, duplicated messages): 
  multicore interconnect networks often are
\end{description}

@@

We now make up a machine model for evaluation. This model is based on
a 8x4x1 AMD Shanghai machine. The topology of such a machine is shown
in Figure~\ref{fig:gruyere}.

@@

    % top horizontal
    \draw[thick] (c1) -- (c2);
    \draw[thick] (c2) -- (c3);
    \draw[thick] (c3) -- (c4);
    % bottom horizontal
    \draw[thick] (c5) -- (c6);
    \draw[thick] (c6) -- (c7);
    \draw[thick] (c7) -- (c8);
    % vertical
    \draw[thick] (c1) -- (c5);
    \draw[thick] (c2) -- (c6);
    \draw[thick] (c3) -- (c7);
    \draw[thick] (c4) -- (c8);
    % cross
    \draw[thick] (c3) -- (c8);
    \draw[thick] (c4) -- (c7);
  \end{tikzpicture}
  \caption{Interconnect topology of 8x4x1 AMD Shanghai machine}
  \label{fig:gruyere}
\end{figure}

@@

@@

% Our implementation
Assuming the tree, we now implement an atomic broadcast for multicore
machines. Our implementation is based on the Barrelfish UMP
interconnect driver. It provides reliable channels with in-order
delivery. Flow control is build in for reliability. Communication
starts at the root node, which acts as a
sequentializer. See~\ref{algo:ab} for details.

@@

\newpage
\input{measurements/gruyere_topologies}
\input{measurements/ziger1_topologies}
\input{measurements/sbrinz1_topologies}
\input{measurements/appenzeller_topologies}

>>>>

\documentclass{article}
\usepackage{url,color,xspace,verbatim,subfig,ctable,multirow,listings}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{rotating}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{graphics}
\usepackage{enumitem}
\usepackage{times}
\usepackage{amssymb}
\usepackage[colorlinks=true]{hyperref}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[toc,page]{appendix}
\usepackage{fancyhdr}
% ==================================================

@@

% tikz stuff
\usepackage{tikz}
\usepackage{pgfplots}

% configuration
\usetikzlibrary{shapes,positioning,calc,snakes,arrows,shapes,fit,backgrounds}
\pgfplotsset{width=.9\linewidth}
\usepgfplotslibrary{external}
%\tikzexternalize

@@

\fancyhf{}
\fancyfoot[L]{\today}
\pagestyle{fancy}

% machines
\newcommand{\ziger}{Ziger (AMD Istanbul 4x6x1)}

@@

\begin{description}
\item[cost breakdown] (quantitative) of message send cost:
  dominating propagation time vs.\ dominating send and receive time on
  multicores (per-node). Our model is currently limited to only the
  breakdown of the cost for sending messages. Maybe we do not strictly
  need to know about propagation time, as it is negligible independent
  of were messages are sent. Relevant seems to be the number of hops
  rather than the latency expressed in a concrete time.
  \begin{description}
  \item[send time] We assume this independent of which node we
    communicate with (which is true for most of our machines when
    using sender-side buffering)
  \item[receive] On shared memory machines, receiving messages is
    realized by reading a memory location. When the sender writes the
    message, the receiver's copy of the cache-line is invalid. The
    next read on it causes a transfer of the cache line from the
    sender's to the receiver's cache, which seems to have a cost that
    is growing with the number of hops on the interconnect for our
    machines (although, on some machines, this is only roughly true:
    communication with different NUMA nodes at the same interconnect
    hop distance varies a bit. We ignore this fact for now.)
  \end{description}
\item[bounded delivery]
\item[reliability] (loss, link failure, in-order delivery, phantoms,
  corruption, duplicated messages): 
  multicore interconnect networks often are
\end{description}

@@

We now make up a machine model for evaluation. This model is based on
a 8x4x1 AMD Barcelona machine. The topology of such a machine is shown
in Figure~\ref{fig:gruyere}.

@@

    % top horizontal
    \draw[thick] (c1) -- (c2);
    \draw[thick] (c2) -- (c3);
    \draw[thick] (c3) -- (c4);
    % bottom horizontal
    \draw[thick] (c5) -- (c6);
    \draw[thick] (c6) -- (c7);
    \draw[thick] (c7) -- (c8);
    % vertical
    \draw[thick] (c1) -- (c5);
    \draw[thick] (c2) -- (c6);
    \draw[thick] (c3) -- (c7);
    \draw[thick] (c4) -- (c8);
    % cross
    \draw[thick] (c3) -- (c8);
    \draw[thick] (c4) -- (c7);
  \end{tikzpicture}
  \caption{Interconnect topology of 8x4x1 AMD Barcelona machine}
  \label{fig:gruyere}
\end{figure}

@@

%--------------------------------------------------
\subsubsection{Adaptive Tree}

\stefan{Describe how this works and how we pick links}

@@

% Our implementation
Assuming the tree, we now implement an atomic broadcast for multicore
machines. Our implementation is based on the Barrelfish UMP
interconnect driver. It provides reliable channels with in-order
delivery and sender-side buffering \stefan{Verify this}. Flow control
is build in for reliability. Communication starts at the root node,
which acts as a sequentializer. See~\ref{algo:ab} for details.

@@

\section{Shared memory performance}

Plot~\ref{pgfplot:201306191213} shows some measurements for using a
shared memory FIFO queue for communication. Core 0 sends message by
enqueuing them to the FIFO, while all other cores poll on the queue
for new messages.

The measurement assumes the tsc clocks to be synchronized. The queues
are also infinite. Messages are never garbage-collected.

\begin{figure}
  \caption{Shared memory FIFO performance on brie}
  \label{pgfplot:201306191213}
  \begin{tikzpicture}
    \begin{axis}[
      ybar,
      error bars/y dir=both,%
      error bars/y explicit,%
      xlabel=core,
      xtick={1,2,3},
      ylabel={cost [cycles]}]

    \addplot coordinates {
      (1,488.6) +- (32.1,32.1)
      (2,761.7) +- (124.1,124.1)
      (3,862.5) +- (118.3,118.3)
    };
    \end{axis}
  \end{tikzpicture}
\end{figure}

Plot~\ref{pgfplot:201306191238} shows the same experiment for tomme.

\begin{figure}
  \caption{Shared memory FIFO performance on tomme}
  \label{pgfplot:201306191238}
  \begin{tikzpicture}
    \begin{axis}[
      ybar,
      error bars/y dir=both,
      error bars/y explicit,
      xlabel=core,
      ylabel={Cost [cycles]}]
    \addplot coordinates {
      (1,417.6) +- (71.0,71.0)
      (2,776.2) +- (110.4,110.4)
      (3,774.6) +- (111.9,111.9)
      (4,648.0) +- (175.9,175.9)
      (5,551.2) +- (257.5,257.5)
      (6,652.6) +- (179.2,179.2)
      (7,688.6) +- (116.3,116.3)
      (8,691.2) +- (133.0,133.0)
      (9,721.6) +- (55.8,55.8)
      (10,633.8) +- (135.0,135.0)
      (11,673.8) +- (86.4,86.4)
      (12,705.0) +- (81.8,81.8)
      (13,612.6) +- (199.1,199.1)
      (14,699.2) +- (90.3,90.3)
      (15,691.6) +- (93.0,93.0)
    };
    \end{axis}
  \end{tikzpicture}
\end{figure}

Plot~\ref{pgfplot:201306191603} shows it for gruyere. These do not
make sense. The numbers are far too high, especially since gruyere is
supposed to be synchronized.

This suggest that my implementation is broken ..

\begin{figure}
  \caption{Plot}
  \label{pgfplot:201306191603}
  \begin{tikzpicture}
    \begin{axis}[
      ybar,
      error bars/y dir=both,
      error bars/y explicit,
      xlabel=FIXME,
      ylabel={Cost [cycles]}]
    \addplot coordinates {
      (1,2399.4) +- (3585.1,3585.1)
      (2,8924.9) +- (11981.8,11981.8)
      (3,4922.3) +- (4685.1,4685.1)
      (4,6638.2) +- (3521.0,3521.0)
      (5,5608.7) +- (4185.6,4185.6)
      (6,5337.5) +- (3846.6,3846.6)
      (7,5844.6) +- (2836.9,2836.9)
      (8,5089.0) +- (3135.3,3135.3)
      (9,5264.0) +- (3919.9,3919.9)
      (10,5711.7) +- (3793.0,3793.0)
      (11,5266.9) +- (3712.1,3712.1)
      (12,8197.6) +- (3819.3,3819.3)
      (13,8668.6) +- (3645.8,3645.8)
      (14,8923.8) +- (2964.8,2964.8)
      (15,7195.3) +- (2407.2,2407.2)
      (16,8176.2) +- (3185.2,3185.2)
      (17,8770.7) +- (3001.8,3001.8)
      (18,4171.5) +- (2011.4,2011.4)
      (19,8453.5) +- (3280.2,3280.2)
      (20,13695.1) +- (3137.6,3137.6)
      (21,13806.8) +- (2914.0,2914.0)
      (22,13692.4) +- (2730.3,2730.3)
      (23,14435.9) +- (2983.7,2983.7)
      (24,9797.6) +- (3821.4,3821.4)
      (25,8955.7) +- (4364.6,4364.6)
      (26,7929.3) +- (3493.3,3493.3)
      (27,11541.9) +- (3076.2,3076.2)
      (28,10851.3) +- (4241.4,4241.4)
      (29,8736.3) +- (3280.4,3280.4)
      (30,11648.2) +- (2969.7,2969.7)
      (31,7821.5) +- (2477.0,2477.0)
    };
    \end{axis}
  \end{tikzpicture}
\end{figure}

Plot~\ref{pgfplot:201306191624} shows the same with more
iterations. The result is not more stable than the one before,
though.

\begin{figure}
  \caption{Plot}
  \label{pgfplot:201306191624}
  \begin{tikzpicture}
    \begin{axis}[
      ybar,
      error bars/y dir=both,
      error bars/y explicit,
      xlabel=FIXME,
      ylabel={Cost [cycles]}]
    \addplot coordinates {
      (1,11197.7) +- (8150.8,8150.8)
      (2,10971.4) +- (6418.6,6418.6)
      (3,10364.3) +- (7409.8,7409.8)
      (4,9460.5) +- (6846.3,6846.3)
      (5,11629.7) +- (7938.9,7938.9)
      (6,12266.8) +- (8895.5,8895.5)
      (7,12313.0) +- (8030.9,8030.9)
      (8,13079.3) +- (7031.2,7031.2)
      (9,10919.1) +- (6272.3,6272.3)
      (10,4758.1) +- (3364.5,3364.5)
      (11,4093.8) +- (2121.7,2121.7)
      (12,10077.3) +- (1152.3,1152.3)
      (13,8632.7) +- (1601.4,1601.4)
      (14,6381.6) +- (1607.5,1607.5)
      (15,8215.2) +- (1799.2,1799.2)
      (16,6973.4) +- (1713.0,1713.0)
      (17,8631.0) +- (1750.3,1750.3)
      (18,13615.4) +- (2873.1,2873.1)
      (19,11311.9) +- (1996.6,1996.6)
      (20,17855.8) +- (2420.9,2420.9)
      (21,15193.0) +- (1923.7,1923.7)
      (22,18920.0) +- (2393.1,2393.1)
      (23,19870.6) +- (2618.8,2618.8)
      (24,17947.8) +- (4379.8,4379.8)
      (25,21783.5) +- (4202.5,4202.5)
      (26,24329.6) +- (3695.0,3695.0)
      (27,22965.0) +- (3667.8,3667.8)
      (28,20440.8) +- (4380.6,4380.6)
      (29,17405.7) +- (4385.3,4385.3)
      (30,19202.7) +- (4509.1,4509.1)
      (31,21855.2) +- (3850.9,3850.9)
    };
    \end{axis}
  \end{tikzpicture}
\end{figure}

\newpage
\input{measurements/gruyere_topologies}

Table~\ref{tab:ziger1} shows an overview of topologies for atomic
broadcast on \ziger. We compare these numbers to cost-estimations from
our simulator that incorporates detailed UMP breakdown measurements
done on the same machine. Our estimations are close to what we observe
when running on real hardware. In particular, the order of efficiency
of the topologies matches reality.

The only exception is sequential execution. Our simulator predicts
much worse performance than actually given by hardware. We believe the
cause for this could be hardware extensions (e.g.\ hardware broadcast
units or snooping).

What is particularly interesting is that a badly chosen tree with lots
of communication across NUMA nodes is actually worse than sequentially
sending messages.

\input{measurements/ziger1_topologies}
\input{measurements/sbrinz_topologies}
\input{measurements/appenzeller_topologies}
